---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TCRconvertR

<!-- badges: start -->
[![codecov](https://codecov.io/gh/seshadrilab/tcrconvertr/graph/badge.svg?token=JVURVQO10D)](https://codecov.io/gh/seshadrilab/tcrconvertr)
[![R-CMD-check](https://github.com/seshadrilab/tcrconvertr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/seshadrilab/tcrconvertr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

> **Warning**: This project is in **beta stage**. It is under active development and may be unstable.

**Rename T-cell receptor genes between 10X, Adaptive, and IMGT formats**

TCRconvertR takes T-cell receptor (TCR) data containing V, D, J, and/or C genes from 10X, Adaptive, or other sequencing platforms and renames them from any of these formats to any other one:

* **10X**: TRAV1-2
* **Adaptive**: TCRAV01-02*01
* **IMGT**: TRAV1-2*01

TCRconvertR works with human, mouse, and rhesus macaque data out-of-the-box, but users can also add their own species.

TCRconvertR helps researchers unify TCR datasets by converting them to a standard naming convention. It is fast, reliable, and prevents errors from manual conversions. Unlike other tools that require custom objects, TCRconvertR works directly with DataFrames and CSV/TSV files.

# Installation

You can install TCRconvertR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("seshadrilab/tcrconvertr")
```

# Usage

**Load some 10X data**

```{r example}
library(TCRconvertR)

tcr_file <- "/home/emmabishop/workspace/tcrconvertr/inst/extdata/examples/tenx.csv"

tcrs <- read.csv(tcr_file)
tcrs
```

**Convert gene names from the 10X format to the Adaptive format**

```{r}
new_tcrs <- convert_gene(tcrs, frm = "tenx", to = "adaptive")
new_tcrs
```

# Contributing

I welcome feedback! If you would like to resolve an issue or add improvements please submit a pull request.

# Issues

If you run into problems or need help running TCRconvertR please file an issue on GitHub.

# Contact

For other questions please contact Emma Bishop: `emmab5` at `uw` dot `edu`

# Acknowledgments

This project was created with support from the Fred Hutchinson Cancer Center Translational Data Science Integrated Research Center (TDS IRC) through the 2024 Data Scientist Collaboration Grant.
