---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![codecov](https://codecov.io/gh/seshadrilab/tcrconvertr/graph/badge.svg?token=JVURVQO10D)](https://codecov.io/gh/seshadrilab/tcrconvertr)
[![R-CMD-check](https://github.com/seshadrilab/tcrconvertr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/seshadrilab/tcrconvertr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Convert T cell receptor (TCR) gene names

`TCRconvertR` converts V, D, J, and/or C gene names between the 10X Genomics, Adaptive Biotechnologies, and IMGT nomenclatures. It supports alpha-beta and gamma-delta TCRs for human, mouse, and rhesus macaque. Users can also define custom species (see docs). A Python version with command-line support is also available.

## Background

TCR annotation tools use different gene naming conventions, making cross-dataset searches difficult (e.g., identifying 10X-annotated TCRs in Adaptive data). Manual conversion is complex and error-prone due to inconsistencies in naming rules.

`TCRconvert` automates this process efficiently and accurately. Our approach is based on analyzing multiple 10X and Adaptive data sets to capture their naming variations. Additionally, we provide a guide (see docs) for generating the necessary gene name columns in Adaptive data when missing.

## Installation

Install from [GitHub](https://github.com/):

```r
# install.packages("pak")
pak::pak("seshadrilab/tcrconvertr")
```

## Usage

**1. Load your TCRs into a data frame**

Example input files:

- **10X**: `filtered_contig_annotations.csv`
- **Adaptive**: `Sample_TCRB.tsv`
- **IMGT**: Output from `MiXCR` or other tools

```{r}
library(TCRconvertR)

tcr_file <- get_example_path("tenx.csv")
tcrs <- read.csv(tcr_file)[c("barcode", "v_gene", "d_gene", "j_gene", "c_gene", "cdr3")]
tcrs
```

**2. Convert**

```{r}
new_tcrs <- convert_gene(tcrs, frm = "tenx", to = "adaptive")
new_tcrs
```

## Contributing

Feedback is welcome! To contribute, submit a pull request.

## Issues

For problems or questions, file an issue on GitHub.

## Contact

For other inquiries, contact Emma Bishop: emmab5 at uw dot edu.

## Acknowledgments

This project was supported by the Fred Hutchinson Cancer Center Translational Data Science Integrated Research Center (TDS IRC) through the 2024 Data Scientist Collaboration Grant. Special thanks to Scott Chamberlain for development support and Shashidhar Ravishankar for gene name curation.
