---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

Install from GitHub:

``` r
# install.packages("pak")
pak::pak("seshadrilab/tcrconvertr")
```

## Basic usage

#### 1. Load TCRs into a data frame

Examples of files you may want to load:

- **10X**: `filtered_contig_annotations.csv`
- **Adaptive**: `Sample_TCRB.tsv`
- **IMGT**: Output from `MiXCR` or other tools

```{r}
library(TCRconvertR)

tcr_file <- get_example_path("tenx.csv") # Using built-in example file
tcrs <- read.csv(tcr_file)[c("barcode", "v_gene", "j_gene", "cdr3")]
tcrs
```

####  2. Convert

```{r}
new_tcrs <- convert_gene(tcrs, frm = "tenx", to = "adaptive")
new_tcrs
```

> Tip: Suppress messages by setting `verbose = FALSE`.

> Tip: If your Adaptive data lacks `x_resolved`/`xMaxResolved` columns, create them yourself by combining `x_gene`/`xGeneName` and `x_allele`/`xGeneAllele` data using `*` as a seperator.


## AIRR data

Supply the standard AIRR gene column names to `frm_cols`:

```r
new_airr <- convert_gene(airr, frm = "imgt", to = "adaptive", 
                         frm_cols = c('v_call', 'd_call', 'j_call', 'c_call'))
```

## Custom column names

By default, `TCRconvertR` assumes these column names based on the input nomenclature (`frm`):

- `frm = 'imgt'` : `c('v_gene', 'd_gene', 'j_gene', 'c_gene')`
- `frm = 'tenx'` : `c('v_gene', 'd_gene', 'j_gene', 'c_gene')`
- `frm = 'adaptive'` : `c('v_resolved', 'd_resolved', 'j_resolved')`
- `frm = 'adaptivev2'` : `c('vMaxResolved', 'dMaxResolved', 'jMaxResolved')`

You can override these columns using `frm_cols`:

**1. Load 10X data with custom column names**

```{r}
custom_file <- get_example_path("customcols.csv")

custom <- read.csv(custom_file)
custom
```

**2. Specify names using `frm_cols` and convert to IMGT**

```{r}
custom_new <- convert_gene(
  custom,
  frm = "tenx",
  to = "imgt",
  verbose = FALSE,
  frm_cols = c("myVgene", "myDgene", "myJgene", "myCgene"),
)
custom_new
```

## Rhesus or mouse data

Use `species = "rhesus"` or `species = "mouse"`

```{r}
new_tcrs <- convert_gene(
  tcrs,
  frm = "tenx",
  to = "imgt",
  species = "rhesus", # or 'mouse'
  verbose = FALSE
)
new_tcrs
```


